Supplementary file

### 2022-11-22 manuscript/manuscript.Rmd
add figure 3 and its code, and new function
- New function to calculate logistic GAMs providing a numeric response and a threshold
- Figure 3 and its code
several updates
- add cutoff scores for delay in developmental domains
- add report functionalities for binomial family of gam regressions
- update manuscript with logistic regressions sections

**Aim**: To describe and model the relationship between sociodemographics, prematurity and neurodevelopmental levels based on the *Ages and Stages Questionnaire* (ASQ-3) scores in infants diagnosed with congenital hypotonia (CH). **Method**: A total of 234 patients diagnosed with CH participated in this study (median age, 8 years; 94 (40.2%) females and 140 (59.8%) males). Neurodevelopmental status was assessed with the ASQ-3 at admission, as well as sociodemographic and obstetric data obtained through the initial clinical interview. **Results**: When modelling the neurodevelopmental status of each domain, an overall negative effect of corrected age on communication and problem-solving skills was observed, whereas the overall effect tends to be positive in gross motor function. Fine motor skills did not exhibit a linear relationship with corrected age but a non-linear effect, unlike the personal-social domain, which did not present significant variations across age corrected for prematurity. **Interpretation**: After adjusting for possible confounders, we found between-subjects fluctuations in neurodevelopmental traits across age in hypotonic infants. These fluctuations were present in the form of non-linear and domain-specific variations.

@@

   - Describes levels of variations observed in neurodevelopmental traits are present across age in hypotonic infants.
   - Observed variations in developmental domains are not solely attributable to prematurity, where age corrected for prematurity best explains the observed variability in neurodevelopment.

@@

In order to assess the influence of confounding variables, we fitted different models using the variables: i) sex of the infant, ii) clinician ID and iii) the relationship of the infant to the respondent (i.e., whether the respondent was the infant's mother, father, uncle or grandparent) as random effects in the form of penalized parametric terms [@wood2016smoothing]. Varying random effects structures models can be seen in the *random effects*' section of the supplementary file.

@@

When modelling the effect of chronological age on developmental domains, different random effects structures were compared when assessing model estimates, these estimates can be seen in [Figure 1](#fig1).

@@

@@

\{Add table\}

@@

>>>>

**Aim**: To describe and model the relationship between sociodemographics, prematurity and neurodevelopmental levels based on the *Ages and Stages Questionnaire* (ASQ-3) scores in infants diagnosed with congenital hypotonia (CH). **Method**: A total of 234 patients diagnosed with CH participated in this study (median age, 8 years; 94 (40.2%) females and 140 (59.8%) males). Neurodevelopmental status was assessed with the ASQ-3 at admission, as well as sociodemographic and obstetric data obtained through the initial clinical interview. **Results**: When modeling the neurodevelopmental status of each domain, an overall negative effect of corrected age on communication and problem-solving skills was observed, whereas the overall effect tends to be positive in gross motor function. Fine motor skills did not exhibit a linear relationship with corrected age but a non-linear effect, unlike the personal-social domain, which did not present significant variations across age corrected for prematurity. **Interpretation**: After adjusting for possible confounders, we found between-subjects fluctuations in neurodevelopmental traits across age in hypotonic infants. These fluctuations were present in the form of non-linear and domain-specific variations.

@@

   - Describes the levels of variations observed in neurodevelopmental traits across age in hypotonic infants.
   - Provides the estimated probability of developmental delay according to the ASQ-3 across age in hypotonic infants.
   - The variations observed in developmental domains are not solely attributable to prematurity, where age corrected for prematurity best explains the variability observed in neurodevelopment.

@@

In order to assess the influence of confounding variables, we fitted different models using the variables: i) sex of the infant, ii) clinician ID and iii) the relationship of the infant to the respondent (i.e., whether the respondent was the infant's mother, father, uncle or grandparent) as random effects in the form of penalized parametric terms [@wood2016smoothing]. Varying random effects structures models can be seen in the section *Random effects structures* of the supplementary file.

Based on the technical report of the ASQ-3, we've been able to estimate the threshold of possible delay based on the total score for each developmental domain. In this sense, a logistic regression, within GAM framework, was used to estimate the predicted probability of having a possible delay in developmental domains given the corrected age.

@@

When modeling the effect of chronological age on developmental domains, different random effects structures were compared when assessing model estimates, these estimates can be seen in [Figure 1](#fig1).

@@

```{r, eval=FALSE}
ind <- grep("interpretacion", names(dataset), value = TRUE)
names(ind) <- c("CM", "GM", "FM", "CG", "PS")

models <- mapply(
  FUN = gam_ordinal,
  var = ind,
  data = list(dataset),
  plot = FALSE,
  SIMPLIFY = FALSE,
  USE.NAMES = TRUE
)

lapply(models, summary)
lapply(models, report_smooth)
lapply(models, report_slopes, term = "edad_corregida_meses")
```

When analyzing the probability of having a developmental delay based on the scores for each domain, we found a significant non-linear effect for CM ($\chi^2_{smooth}$ (7.73) = 32.11, *p* < 0.001), whereas in this domain we observed that for every one-month increase in the corrected age in the 13.6 to 18.9 age range, the odds of a possible developmental delay increases by 60% (OR = 1.6 CI~95%~[1.16, 2.23], $t_{student}$ (224.56) = 2.89, *p* = 0.01), as well with the 28.6 to 29.6 age range, whereas the odds of delay increases by 36% in this domain (OR = 1.36 CI~95%~[1.00, 1.86], $t_{student}$ (224.56) = 1.98, *p* = 0.049). Similar to CM, an overall linear effect of corrected age in the GM domain was observed ($\chi^2_{smooth}$ (1) = 6.86, *p* = 0.009), whereas for every one-month increase the odds of a possible developmental delay decrease by a 3.9% (OR = 0.96, CI~95%~[0.50, 0.99], $t_{student}$ (232) = -2.62, *p* = 0.009). The probability associated with a possible delay according to the total score for each developmental domain in response to corrected age can be observed in [Figure 3](#fig3).

@@

Añadir tabla

@@

[**Figure 3**]{#fig3}. Predicted probability of having a possible delay according to the total score for each developmental domain in response to corrected age. Fitted models were estimated using logistic regression within GAM framework and adjusted for possible confounders as already described. The p-values of smooth terms, associated with the main effect of corrected age on developmental delay, are shown within each panel.
